---
title: "Homework 4"
author: "Erika Luna"
date: "October 6, 2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
library(dplyr)
library(ggplot2)
library(knitr)
library(DT)
library(devtools)
```

#Exercise 1: Univariate Data Reshaping

##Univariate option 1
I've chose this option, for which I will need to:

1.1 Make a tibble with one row per year, and columns for life expectancy for two or more countries.
1.2 Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.
1.3 Re-lengthen the data.

##1.1 

```{r}
dat = gapminder %>%
  filter(country == "Mexico" | country == "Canada") %>%
  select(country, year, lifeExp) %>%
  pivot_wider(id_cols = year,
              names_from = country,
              names_prefix = "lifeExp_",
              values_from = lifeExp)
DT::datatable(dat)
```

##1.2
```{r}
axis_lim = c(min(dat$lifeExp_Mexico, dat$lifeExp_Canada),
             max(dat$lifeExp_Mexico , dat$lifeExp_Canada))
dat %>%
  ggplot(aes(lifeExp_Mexico, lifeExp_Canada)) +
  geom_point(aes(colour = "red"), size = 3) +
  geom_abline(linetype = 2, alpha = 0.3) +
  geom_text(aes(label = year), size = 3, hjust = 1.3) +
  coord_fixed(xlim = axis_lim, ylim = axis_lim) +
  labs(title = "Life Expectancy (years) of Canada against Mexico",
       x = "Life Expectancy in Mexico",
       y = "Life Expectancy in Canada") +
  theme_bw() +
  theme(legend.position = "none")
```

##1.3
```{r}
dat %>%
  pivot_longer(cols = c(lifeExp_Mexico, lifeExp_Canada),
               names_to = "country",
               names_prefix = "lifeExp_",
               values_to = "lifeExp") %>%
  DT::datatable()
```

#Exercise 2: Multivariate Data Reshaping

##Multivariate Option 1: 
I've chosen this option, for which I will need to:

2.1 Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two or more countries.

2.2 Re-lengthen the data.


##2.1
```{r}
dat = gapminder %>%
  filter(country == "Mexico" | country == "Canada") %>%
  select(country, year, gdpPercap, lifeExp) %>%
  pivot_wider(id_cols = year,
              names_from = country,
              names_sep = "_",
              values_from = c("lifeExp", "gdpPercap"))
DT::datatable(dat)
```

##2.2

```{r}
dat %>%
  pivot_longer(cols = -year,
               names_to = c(".value" ,"country"),
               names_sep = "_",
               values_to = c("lifeExp", "gdpPercap")) %>%
  DT::datatable()
```

#Exercise 3: Table Joins

For this exercise wee need to load the following datasets: 

```{r echo = FALSE, warning = FALSE, message = FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

##3.1

For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble.


```{r}
email_modified <- 
  separate_rows(email, guest, sep = ", " ) %>% 
  rename(name = guest)

guest %>% 
  left_join(email_modified, by = "name") %>% 
  DT::datatable(options = list(scrollX = TRUE))
```

##3.2

Who do we have emails for, yet are not on the guestlist?

```{r}
non_guest <- email_modified %>%
  anti_join(guest, by = "name")
  knitr::kable(non_guest)
```

##3.3

Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r}
guest %>% 
  full_join(email_modified, by = "name") %>%
  DT::datatable()
```