---
title: "Homework Assignment 2"
author: "Erika Luna"
date: "22/09/2019"
output:
  html_document:
    keep_md: yes
    number_sections: no
    toc: no
    toc_float: yes
  pdf_document:
    toc: no
---

```{r load, warning=FALSE, echo=FALSE}
library(gapminder)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

## Exercise 1

## 1.1

```{r}
gapminder_modified <- filter(gapminder, year < 1980 & year > 1970 &
       country == "Brazil" |
         country == "Israel" |
         country == "Portugal")
```

## 1.2

```{r}
gapminder_modified %>% 
  select(country, gdpPercap)
```

## 1.3

```{r}
gapminder %>% 
  mutate(lifeExpInc = lifeExp - lag(lifeExp)) %>% 
  filter(lifeExpInc > 0)

```

## 1.4

```{r}
gapminder %>% 
       group_by(country) %>% 
        filter(gdpPercap == max(gdpPercap))
```

## 1.5

```{r}
gapminder %>% 
  filter(country == 'Canada') %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point() +
  scale_x_log10()
```


## Exercise 2

Categorical variable: Continet

There are 5 possible values for the variable Continent:

```{r}
summary(gapminder$continent)
```

The following graphic shows the amount of data per continent:

```{r}
ggplot(gapminder, aes(continent)) +
  geom_bar()+
  ylab("Count")+
  xlab("Continent")
```

Quantitative variable: Population

```{r}
range(gapminder$pop)
```

The following graphic shows the ammount of population in each continent:

```{r}
ggplot(gapminder, aes(continent, pop)) +
  geom_col()
```






Then we can also look at the distribution for life expectancy over the continents:
```{r}
ggplot(gapminder, aes(continent, gdpPercap)) +
  #geom_point() +
  geom_boxplot()
```


## Exercise 3

```{r}
library(datasets)
```

I will use the 'quakes' data frame which has information about earthquakes off Fiji:

```{r}
head(quakes)
```

Based on the following plot we can say that there is not a clear relationship between the depth and the magnitude of earthquakes:

```{r}
quakes %>% 
  ggplot(aes(depth, mag)) +
  geom_point() +
  xlab('Depth') +
  ylab('Magnitude') +
  ggtitle('Magnitude of earthquakes in Fiji')
```

Based on the following graph we can see which stations registered the most earthquakes. A spatial analysis to locate the stations could also be useful for our analysis. 

```{r}
quakes %>% 
  ggplot(aes(stations)) +
  geom_bar(aes()) +
  xlab('Stations') +
  ylab('Number of earthquakes')+
  ggtitle('Earthquakes registered at each station')
```


## Recycling

For people who want to take things further.

Evaluate this code and describe the result. Presumably the analyst???s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?


```{r}
DT::datatable(filter(gapminder, country == c("Rwanda", "Afghanistan")))
```

The code above produces a table with data for Afganistan and Rwanda only; therefore, it works.
