---
title: "Homework 5"
author: "Erika Luna"
date: "October 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
library(dplyr)
library(DT)
library(knitr)
library(here)
library(tsibble)
library(cowplot)
library(ggplot2)
knitr::opts_chunk$set(fig.width=4, fig.height=3, warning = FALSE, fig.align = "center")
#knitr::opts_chunk$set(echo = TRUE)
```


# Exercise 1: Explain the value of the here::here package
```{r}

```



# Exercise 2: Factor management

Task: Choose one dataset (of your choice) and a variable to explore. After ensuring the variable(s) you are exploring are indeed factors, you should:

```{r}
str(gapminder$continent)
```


## 2.1: Drop factor / levels

```{r, warning = FALSE, message = FALSE}
gapminder$continent %>% 
  nlevels() 
```

```{r}
nrow(gapminder)
```


```{r}
gapminder_dropped <- gapminder %>% 
  filter(continent == c("Asia", "Europe", "Americas", "Africa")) %>% 
  droplevels()
```

```{r}
gapminder_dropped$continent %>% 
  nlevels()
```
```{r}
gapminder_dropped %>% 
  nrow()
```



## 2.2 Reorder levels based on knowledge from data

```{r}
gap_maxlife <- gapminder_dropped %>% 
  group_by(continent, country) %>% 
  summarize(max_lifeExp = max(lifeExp))
```

```{r}
gap_maxlife$country %>% 
  nlevels()
```

```{r}
gap_europe <- gap_maxlife %>% 
  filter(continent == "Europe") %>% 
  droplevels()
```


```{r}
gap_europe %>% 
  ggplot() +
  geom_point(aes(country, max_lifeExp)) +
  ylab("Maximum life expectancy (yeas)") +
  xlab("Country") +
  coord_flip() +
  theme_light()
```



```{r}
gap_europe %>% 
  ggplot() +
  geom_point(aes(fct_reorder(country, max_lifeExp, min), max_lifeExp)) +
  ylab("Maximum life expectancy (years)") +
  xlab("Country") +
  coord_flip() +
  theme_light()
```

# Exercise 3: File input/output (I/O)

Let's say we are interested in sharing the data for America with our collegues. We first have to filter our gapminder data. 
```{r}
gapminder_ame <- gapminder %>% 
  filter(continent == "Americas")
```

Now we write a csv of the filtered data.
```{r}
write_csv(gapminder_ame,
          here::here("hw05", "gapminder_ame.csv"))
```
Now, when our collegues open the file, they would use the function 'read_csv'. 

```{r}
gapminder_ame_csv <- read_csv(here("hw05", "gapminder_ame.csv"))

gapminder_ame_csv %>%
  knitr::kable()
```

```{r}
class(gapminder_ame_csv$country)
class(gapminder_ame_csv$continent)
class(gapminder_ame_csv$year)
class(gapminder_ame_csv$lifeExp)
class(gapminder_ame_csv$pop)
class(gapminder_ame_csv$gdpPercap)
```

```{r}
gapminder_ame_csv %>% 
  mutate(country = fct_reorder(country, pop, .desc = TRUE)) %>% 
  arrange(continent)
```



# Exercise 4

On Assignment 1 I made the graph shown below. At the time it seemed nice to me, but now that I've learned how to make pretty and most importantly, informative graphs with ggplot I will change some aspects if it. 

```{r}
old_plot <- ggplot(gapminder, aes(continent, pop)) + 
  geom_col()
```

```{r}
new_plot <- gapminder %>% 
  ggplot() +
  geom_col(aes(reorder(continent, -pop), pop)) +
  ylab("Population") +
  xlab("Continent") +
  scale_y_continuous(labels = scales::comma) +
  theme_light()
```

```{r}
plots <- plot_grid(old_plot, new_plot, ncol =,
          labels = c("Old plot", "New plot"))

plots
```



# Exercise 5: Writing figures to file

Task: Use ggsave() to explicitly save a plot to file. Include the exported plot as part of your repository and assignment.

```{r}
plots_file <- "plots.png" #file name

plot_path <- here::here("hw05", "plots") #path

ggsave(plots_file, plots, device = "png", path = plot_path, width = 10, height = 4)
```














