---
title: "Homework 3"
author: "Erika Luna"
date: "October 1, 2019"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this assignment we will explore the gapminder dataset using the dplyr package. Before we start we need to load a couple of packages. 


```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(gapminder)
library(dplyr)
library(DT)
library(knitr)
```

I have decided to work on tasks 2, 3 and 6.

## Task Option 2
Instructions: Get the maximum and minimum of GDP per capita for all continents.

We first produce a tibble (a type of data frame).

```{r tibble task2}
task_2 <- gapminder %>% #We name our modified dataset
  group_by(continent) %>% #We use this function in order to obtain the information at the continent level.
  summarise(min_gdp = min(gdpPercap), #then we use this function in order to obtain the maximum and minimum of GDP per capita. 
            max_gdp = max(gdpPercap))
  
task_2
```

From the tibble we produced above, we now create a plot. 

```{r plot task2}
task_2 %>% 
  ggplot(aes())+ #We can plot the GDP max and min in the same plot by using geom_point twice. We could also use some other type of "geom". 
  geom_point(aes(continent,max_gdp, colour = "red"))+ 
  geom_point(aes(continent,min_gdp, colour = "blue"))+
  labs(title="Minimum and maximum GDP of each continet",
               x="Continent",
               y="GDP per capita")+
  scale_color_discrete("",labels=c("Minimum GDP", "Maximum GDP"))+
  scale_y_log10(labels = scales::dollar_format())


```




## Task Option 3

Instructions: Look at the spread of GDP per capita within the continents.

We first create a table. 
```{r tibble task3}
gapminder %>% 
  group_by(continent) %>% 
  summarize(max_gdpPercap = max(gdpPercap),#We can use a couple of different functions within the summarize function. 
            min_gdpPercap = min(gdpPercap),
            range_gdpPercap = max(gdpPercap)-min(gdpPercap),
            sd_gdpPercap = sd(gdpPercap),
            mean_gdpPercap = mean(gdpPercap),
            median_gdpPercap = median(gdpPercap)) %>% 
  knitr::kable()
```

Now we create a plot.
```{r plot task3}
gapminder %>% 
  ggplot(aes(continent, gdpPercap)) +
  geom_boxplot() + #We use this type of "geom" as we want to show the spread of GDP.
  scale_y_log10(labels = scales::dollar_format()) + #We can add a dollar sign to the values on the y axis.
  labs(title = "Spread of GDP per capita for all continents", 
       x = "Continent", 
       y = "GDP per capita")
```



## Task Option 5

Instructions: How is life expectancy changing over time on different continents?

As before, we first create a table that shows the average life expectancy for each continent for each year. 
```{r tibble task5}
gapminder %>% 
  group_by(continent, year) %>% 
  summarise(mean_life = mean(lifeExp)) %>% 
  DT::datatable()
```

Now we create a plot: 
```{r plot task5}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country, colour = continent)) +
  geom_jitter(size = 1, alpha = 0.5)+ #We use geom_jitter to show the behaviour of each country within each continent. 
  geom_line()+
  facet_wrap(~ continent, nrow = 3)  +
  labs(title = "Change of life expectancy over time", 
       x = "Year", 
       y = "Country life expectancy (mean)")
```
